---
title: "White Wines: A Data Exploration Approach"
author: "Lamont Girton"
date: "November 2, 2015"
output: 
  html_document: 
    theme: journal
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}

library(knitr)

opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.align='center')

library(ggplot2)
library(dplyr)
library(tidyr)

```

```{r}

wines <- read.csv("./wineQualityWhites.csv")

```

# Introduction

This paper will explore the quality of various white wines through their chemical composition, with the purpose of visualizing the data and exploring various variable relationships with respect to _quality_.  The dataset for this paper is available [here](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityWhites.csv).  The provided dataset contains `r nrow(wines)` observations of `r ncol(wines)`.  The variable labeled _X_ is 100% unique and appears to be an instance identifier and sense it provides no explanatory benefit, will be excluded from the analysis in this paper.


# Univariate Analysis



## Dataset Structure
Below is a listing of the _str_ command, which displays the structure of the data frame.  As noted earlier, there are `r ncol(wines)` variables and from the listing below, they all appear to be numeric.  Additionally, there are no missing values.

```{r, echo=TRUE}

str(wines)

```

```{r, results='asis'}
summaries <- data.frame(cbind(
    apply(wines, 2, min),
    apply(wines, 2, mean),
    apply(wines, 2, median),
    apply(wines, 2, max),
    apply(wines, 2, sd)
))

names(summaries) <- c("Min", "Mean", "Median", "Max", "SD")

    
```

## Descriptive Statistics

```{r}
kable(summaries, digits = 3)

```


```{r}

data.frame(cbind(scale(wines[2:12]),quality=wines$quality)) %>% 
    gather(variable, value, -quality) %>% 
    ggplot(aes(value, fill=factor(quality))) +
    geom_histogram() +
    facet_wrap(~variable, scales = "free")
    

```


## Distributions

```{r}

select(wines, -X) %>% 
    gather(variable) %>% 
    ggplot(aes(value)) +
    geom_density() +
    facet_wrap(~variable,scales = "free")

```

# Bivariate Analysis

# Multivariate Analysis

# Final Plots and Summary

# Reflection





